Toward Centrally Programming Distributed Control Planes

### 2016-01-22 paper/compilation.tex
more edits

At a first cut, any time a router must make a decision locally between several route options, there is the possibility that it might choose incorrectly. The product graph representation captures this notion of choice precisely. For example, router $D$ in Figure~\ref{fig:example-compilation} appears only once in the product graph, possibly receiving route advertisements from $Y$ or $X$. However, regardless of whether $D$ chooses a route from $X$ or $Y$, the set of paths allowed by the policy after going through $D$ will be the same in either case. This remains true despite any failures that might have occurred in the network. Thus $D$ can safely prefer $X$ and $Y$ equally.

The more challenging case is when a topology location occurs in mutliple contexts in the product graph. For example, the topology node $C$ can receive an advertisement from $E$ and later achieve the backup path, or it can receive an advertisement from its neighbor $D$ and later achieve either the primary or backup path. Is it safe for $C$ to prefer its neighbor $D$ over its other neighbor $C$? The important observation is that, if $C$ prefers $D$, then it will always achieve at least as good a path as if it had choosen $E$. 
For example, suppose $C$ chooses a route from $D$, but cannot achieve its primary path because the $A$ -- $W$ link has failed. In this case, the advertisement from $C$ will still be sent along towards the ultimate backup location $W$. Since the $(-,2) C$ and $(3,2) C$ nodes in the product graph share the same next hops in either case, no possible failure will prevent a route advertisment from reaching $(-,4) W$ that wouldn't have otherwise prevented it if $C$ had choosen $E$.

In general, ensuring that an individual router's preferences respect the policy under all possible failures is hard. A simple enumeration of failures is intractable. Instead, we adopt a conservative analysis based on the product graph representation, and the previous observations, which runs efficiently in practice. The high-level idea is to order each node in the product graph according to an \textit{can prefer} relation. For example, node $(3,2) C$ can be preferred to node $(-,2) C$, but not the other way around. Intuitively, a node $N_1$ can be preferred to another node $N_2$ when, for each preference $N_2$ might achieve, there is a better preference that $N_1$ will achieve regardless of failures. Formally this means that $N_1$ can be preferred to $N_2$ when:
%
$$\forall i, \exists j, j \leq i \wedge protect(G_j, N_1, G_i, N_2)$$
%
The algorithm listed in Figure~\ref{fig:algorithm} defines what it means for one node to \textit{protect} against the failures of another. Intuitively, the algorithm walks from nodes $N_1$ and $N_2$ in $G_j$ and $G_i$ respectively, and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, take an equivalent step. Thus $N_1$ any paths leading from $N_2$ will have an equivalent path from $N_1$.

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\begin{algorithm}
\caption{Failure Protection}\label{euclid}
\begin{algorithmic}[1]
\Procedure{Protect($G_1$, $N_1$, $G_2$, $N_2$)}{}
\If {$N_1.loc \neq N_2.loc$} \Return false
\EndIf
\State $q \gets Queue()$
\State $q.Push (N_1, N_2)$
\While {\textit{not} $q.Empty()$}
\State TODO
\State TODO
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

TODO: how simple paths factor in

TODO:

>>>>

At a first cut, any time a router must make a decision locally between several route options, there is the possibility that it might choose incorrectly. The product graph representation captures this notion of choice precisely. For example, router $D$ in Figure~\ref{fig:example-compilation} appears only once in the product graph, possibly receiving route advertisements from $X$ or $Y$. However, regardless of whether $D$ chooses a route from $X$ or $Y$, the set of paths allowed by the policy after going through $D$ will be the same in either case. This remains true despite any failures that might have occurred in the network. Thus $D$ can safely prefer $X$ and $Y$ equally.

The more challenging case is when a topology location occurs in mutliple contexts in the product graph. For example, the topology node $C$ can receive an advertisement from $E$ and later achieve the backup path, or it can receive an advertisement from its neighbor $D$ and later achieve either the primary or backup path. Is it safe for $C$ to prefer its neighbor $D$ over its other neighbor $E$? The important observation is that, if $C$ prefers $D$, then it will always achieve at least as good a path as if it had choosen $E$. 
For example, suppose $C$ chooses a route from $D$, but cannot achieve its primary path because the $A$ -- $W$ link has failed. In this case, the advertisement from $C$ will still be sent along towards the ultimate backup location $W$. Since the $(3,2) C$ node has the same one-step and two-step next hops as node $(-,2) C$ in the product graph, no possible failure will prevent a route advertisment from reaching $(-,4) W$ that wouldn't have otherwise prevented it if $C$ had choosen $E$.

In general, ensuring that an individual router's preferences respect the policy under all possible failures is hard. A simple enumeration of failures is intractable. Instead, we adopt a conservative analysis based on the product graph representation and the previous observations, which runs efficiently in practice. The high-level idea is to order each node in the product graph according to a \textit{can prefer} relation. For example, node $(3,2) C$ can be preferred to node $(-,2) C$, but not the other way around. Intuitively, a node $N_1$ can be preferred to another node $N_2$ when, for each preference $N_2$ might achieve, there is a better preference that $N_1$ will achieve regardless of failures. Formally this means that $N_1$ can be preferred to $N_2$ when:
%
$$\forall i, \exists j, j \leq i \wedge protect(G_j, N_1, G_i, N_2)$$
%
where the \textit{protect} predicate means that, from $N_1$ on the product graph restricted to preferences $j$ or better, there are at least as many paths as from $N_2$ restricted to $G_i$

The algorithm listed in Figure~\ref{alg:failures} defines what it means for one node to \textit{protect} against the failures of another. Intuitively, the algorithm walks from nodes $N_1$ and $N_2$ in $G_j$ and $G_i$ respectively, and ensures that for every \textit{step} $N_2$ can take to some new topology location, $N_1$ can, at the very least, take an equivalent step. Thus $N_1$ any paths leading from $N_2$ will have an equivalent path from $N_1$.

\begin{algorithm}[t!]
\caption{Failure Protection}
\label{alg:failures}
\begin{algorithmic}[1]
\Procedure{Protect($G_1$, $N_1$, $G_2$, $N_2$)}{}
  \If {$N_1.loc \neq N_2.loc$} \Return false
  \EndIf
  \State $q \gets Queue()$
  \State $q.Enqueue (N_1, N_2)$
  \While {$!q.Empty()$}
    \State $(n_1,n_2) \gets q.Dequeue()$
    \For {$x$ in $adj(G_2, n_2)$}
      \If {exists a $y$ where $y$ shadows $x$} 
        \If {$(x,y)$ not marked}
          \State {mark $(x,y)$ as seen}
          \State $q.Enqueue(x,y)$
        \EndIf
      \Else { \Return $false$}
      \EndIf
    \EndFor
  \EndWhile
  \Return true
  \EndProcedure
\end{algorithmic}
\end{algorithm}
